SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization

### 2014-05-11 paper/paper.tex
paper: submitted version.

\title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization}
\subtitle{UNPUBLISHED DRAFT}

@@

As shown in the motivating example in Section~\ref{sec:example}, SmartCheck
can shrink more than an order-of-magnitude faster than even the most efficient hand-written
shrink implementations for some properties and programs.  As well, the time
required to shrink scales linearly with the size of the counterexample---other
common shrink implementations often scale in polynomial time or worse.
(Shrinking time is important in real-world uses; see
Section~\ref{sec:experiments} for examples.)

@@

In this section, we motivate in more detail the challenges in shrinking
counterexamples by comparing manual approaches using QuickCheck to SmartCheck.
(We focus on shrinking rather than generalization here since counterexample
generalization is unique to SmartCheck.)  We will show how a small data type and
simple property can result in large counterexamples without any shrinking.  Then
we show the difficulty in designing an efficient shrink implementation manually,
which is required when using QuickCheck; we will show a poor
design before arriving at a ``canonical'' manual solution.  However, even the
canonical solution produces significantly larger counterexamples and takes
significantly longer to produce them than SmartCheck, even though SmartCheck's
shrinking strategy is completely automatic and generic.

@@

The most recent released version of QuickCheck provides a \ttp{genericShrink}
function for algebraic data types, so the user merely has to declare
%
\begin{code}
shrink = genericShrink
\end{code}
%
\noindent
to obtain shrink instances.  (The implementation relies on the same generics
package that SmartCheck does; see Section~\ref{sec:preliminaries}.)

@@

              & \multicolumn{2}{c}{Mean} & \multicolumn{2}{c}{Std. dev.} &
\multicolumn{2}{c}{95\%}\\\hline
              & size & time  & size & time & size & time \\\hline\hline
QC none       & 83 & 0.013   & 22 & 0.014   & 119 & 0.023\\\hline
QC trunc      & 40 & 0.064   & 20 & 0.334   & 84 & 0.153\\\hline
QC tuple      & 10 & 0.031   & 4 &  0.039   & 16 & 0.052\\\hline
QC generic    & &        &        &           & &\\\hline
SmartCheck    & 6 & 0.021s    &  4 & 0.009   & 13 & 0.036\\\hline

@@

The results of our discussion above are summarized in Table~\ref{table:results}
and the corresponding graph in Figure~\ref{fig:graphs}. They show four
approaches to shrinking with QuickCheck: no shrinking (QC none), to provide a
baseline, truncated shrinking (QC trunc) with \ttp{Int16} shrinking disabled, in
which each list is truncated to 10 elements, tuple shrinking (QC tuple), also
with \ttp{Int16} shrinking disabled, and finally shrinking using QuickCheck's
\ttp{genericShrink} (qc generic).  We omit SmallCheck, since as noted, it is not
even feasible for the example with default instances.\footnote{All results
  reported in the paper are with a GHC-7.6.3-compiled program, using \ttp{-O2},
  on a 2.8 GHz 4~core i7 with 16GB memory under OSX v. 10.9.2.}  We graph the
final counterexample size, measured in the total number of \ttp{Int16} values,
and execution time for counterexample discovery and shrinking.
Figure~\ref{fig:graphs} shows how many of the 1000 runs resulted in a final
value of a particular size, fitting the data with a Bezier curve.

In Table~\ref{table:results}, we summarize the graphed values and provide
execution time in seconds as well.  For both execution time and final value
size, we summarize the mean, median, standard deviation, and the results at the
95th percentile.  (While we provide the standard deviations, note that the plots
are not necessarily Gaussian.)

SmartCheck produces consistently small counterexamples and does so
quickly---even smaller and faster than the best hand-tuned solution---all
without writing any problem-specific code.  Note that the ``no shrinking''
example provides a lower-bound on execution time (only including counterexample
discovery time) and an upper-bound on counterexample size.  In this example, the
hand-written ``tuple shrink'' algorithm is only slightly slower and produces
slightly larger results.  In Section~\ref{sec:benchmarks}, we provide additional
benchmarks, but now we turn to the design of SmartCheck.

@@

%% Finally, these performance results for SmartCheck are fairly generalizable to
%% other types and properties, since the algorithm is not dependent on the
%% specific program or property being checked.  However, the results are
%% particularly pronounced for counterexamples that are particularly ``deep'' or
%% ``broad'' in terms of their tree structure.  %% For programs and properties
%% for which small counterexamples are easily found, QuickCheck's shrinking is
%% perfectly acceptable.  Furthermore, user-defined shrink methods may vary
%% significantly based on the property and counterexample type, so it is not
%% always clear which implementation to compare against.
We now describe in detail SmartCheck's algorithm for generic shrinking.

@@

\section{Shrinking Data}\label{sec:shrinking}
In this section, we describe how to efficiently and generically shrink algebraic
data values.  Recall the basic idea behind the \ttp{shrink}
method of the \ttp{Arbitrary} class: generate a list of values, each of which is
smaller than the current counterexample.  Each of the new values generated
may not bear any relationship to the original counterexample other than being
smaller.  SmartCheck pursues an approach that searches for smaller but
structurally similar counterexamples, as we make precise below.

@@

\todo{reference into with QC, no int shrinking}

@@

\todo{how long unanswered?}

@@

Unfortunately, no set of testing benchmarks exists over which to compare
different test-case generation and minimization approaches.  Therefore, we have
collected a small number of benchmarks, in addition to the more involved
case-studies described earlier in this section.

The benchmarks presented, in addition to the motivating example presented in
Section~\ref{sec:example}, compare standard SmartCheck against hand-written
shrink implementations.  The benchmarks are as follows:

@@

  \item \emph{Parser}, a parser/pretty-printer for a toy imperative language
    containing a parser bug that switches the arguments of an binary operator.
\end{itemize}

@@

\todo{generic or hand-written?}

@@

\multirow{2}{*}{Reverse}
  & QC & 2 & 0.002         & 0 & 0.002         & 2 & 0.007 \\\cline{2-8}
  & SC & 2 & \sci{4.8}{-4} & 0 & \sci{8.1}{-4} & 2 & \sci{5.8}{-4}\\\hline\hline

\multirow{2}{*}{Div0}
  & QC & 10 & 0.001 & 6 & 0.002 & 23 & 0.002 \\\cline{2-8}
  & SC & 5 & 0.001  & 0 & 0.001 & 5 & 0.002  \\\hline\hline

\multirow{2}{*}{Heap}
  & QC & 15 & 0.016  & 1 & 0.025 & 19 & 0.045 \\\cline{2-8}
  & SC & 7  & 0.007  & 2 & 0.003 & 12 & 0.015 \\\hline\hline

\multirow{2}{*}{Parser}
  & QC &  191 & 0.013 & 195 & 0.018 & 577 & 0.044 \\\cline{2-8}
  & SC &  16  & 0.241 & 75  & 0.800 & 14  & 0.521 \\\hline

@@

The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit
of using SmartCheck to discover small counterexamples, since a list of length
two falsifies the property (SmartCheck is still useful in generalizing the
counterexample, however!).  Perhaps surprisingly, SmartCheck still slightly
outperforms QuickCheck.  The other two benchmarks have much larger counterexamples,
where SmartCheck's benefits are more pronounced.  For example, consider the
parser example in which the initial counterexample to shrink may be in a large
parse tree.  SmartCheck performs significantly better than optimal hand-written
shrink instances, but even SmartCheck can sometimes return large
counterexamples---this is why the mean counterexample size is larger than the
size at the 95th percentile: there are a small number of large values not
successfully shrunk by SmartCheck.

@@

Other researchers have also investigated counterexample shrinking.  Zeller
and Hildebrandt describe an application of greedy search to shrink
counterexamples they call ``delta-debugging'' (DD)~\cite{dd}.  The authors apply
their work to shrinking HTML inputs to crash Mozilla and shrinking C programs to
trigger a bug in GCC.  Subsequent generalizations are reported by Misherghi and
Su in which they perform greedy search on tree-structured data; they call their
approach hierarchical delta-debugging (HDD)~\cite{hdd}.

>>>>

\title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization}
%\subtitle{UNPUBLISHED DRAFT}

@@

%% As shown in the motivating example in Section~\ref{sec:example}, SmartCheck
%% generates smaller counterexamples faster than hand-written or generated
%% QuickCheck shrinking, for some properties and programs.  (Shrinking time is
%% important in real-world uses; see Section~\ref{sec:experiments} for examples.)

@@

In this section, we motivate in more detail the challenges in shrinking
counterexamples by comparing manual approaches using QuickCheck to SmartCheck.
(We focus on shrinking rather than generalization here since counterexample
generalization is unique to SmartCheck.)  We will show how a small data type and
simple property can result in large counterexamples without any shrinking.  Then
we show the difficulty in designing an efficient shrink implementation manually,
which is required when using QuickCheck; we will show a poor
design before arriving at a ``canonical'' manual solution.%%   However, even the
%% canonical solution produces significantly larger counterexamples and takes
%% significantly longer to produce them than SmartCheck, even though SmartCheck's
%% shrinking strategy is completely automatic and generic.

@@

In early 2014, QuickCheck added the ability to derive \ttp{shrink} instances for
algebraic data type using the same generic package that SmartCheck does; see
Section~\ref{sec:preliminaries}.  (The first release of the SmartCheck software
was mid-2012.)  With this new feature, the user declares
%
\begin{code}
shrink = genericShrink
\end{code}
%
\noindent
to obtain efficient shrink instances.

@@

              & \multicolumn{2}{c}{Mean} & \multicolumn{2}{c}{Std. dev.} &
\multicolumn{2}{c}{95\%}\\\hline
              & size & time  & size & time & size & time \\\hline\hline
none          & 83 & 0.013   & 22 & 0.014   & 119 & 0.023\\\hline
QC trunc      & 40 & 0.064   & 20 & 0.334   & 84 & 0.153\\\hline
QC tuple      & 10 & 0.031   & 4  & 0.039   & 16 & 0.052\\\hline
QC generic    & 13 & 0.204   & 7  & 0.946   & 28 & 0.543\\\hline
SmartCheck    & 6 & 0.021s   & 4  & 0.009   & 13 & 0.036\\\hline

@@

The results of our discussion above are summarized in Table~\ref{table:results}
and the corresponding graph in Figure~\ref{fig:graphs}.  First, we show the
results of no shrinking (labeled `none') as a baseline in terms of worst-case
value size.  Then we show four approaches to shrinking, as described above:
truncated shrinking (`QC trunc') with \ttp{Int16} shrinking disabled, in which
each list is truncated to 10 elements, tuple shrinking (`QC tuple'), also with
\ttp{Int16} shrinking disabled, and finally shrinking using QuickCheck's
\ttp{genericShrink} (`QC generic').\footnote{All results
  reported in the paper are with a GHC-7.6.3-compiled program, using \ttp{-O2},
  on a 2.8 GHz 4~core i7 with 16GB memory under OSX v. 10.9.2.}  We graph the
final counterexample size, measured in the total number of \ttp{Int16} values,
and execution time for counterexample discovery and shrinking.
Figure~\ref{fig:graphs} shows how many of the 1000 runs resulted in a final
value of a particular size, fitting the data with a Bezier curve.

In Table~\ref{table:results}, we summarize the graphed values and provide
execution time (including both counterexample discovery and shrinking) in
seconds as well.  For both execution time and final value size, we summarize the
mean, median, standard deviation, and the results at the 95th percentile.
(While we provide the standard deviations, note that the plots are not
necessarily Gaussian.)

SmartCheck produces consistently small counterexamples and does so
quickly---even smaller and faster than the best hand-tuned solution---all
without writing any problem-specific code.  Note that the `none' example
provides a lower-bound on execution time (only including counterexample
discovery time) and an upper-bound on counterexample size.  In this example, the
hand-written ``tuple shrink'' algorithm is only slightly slower and produces
slightly larger results than SmartCheck.  (QuickCheck's generic shrinking under
performs the ``tuple shrink'' likely because the generic shrink implementation
shrinks the \ttp{Int16} values.)

In Section~\ref{sec:benchmarks}, we provide additional benchmarks, but now we
turn to the design of SmartCheck.

@@

%% Finally, these performance results for SmartCheck are fairly generalizable to
%% other types and properties, since the algorithm is not dependent on the
%% specific program or property being checked.  However, the results are
%% particularly pronounced for counterexamples that are particularly ``deep'' or
%% ``broad'' in terms of their tree structure.  %% For programs and properties
%% for which small counterexamples are easily found, QuickCheck's shrinking is
%% perfectly acceptable.  Furthermore, user-defined shrink methods may vary
%% significantly based on the property and counterexample type, so it is not
%% always clear which implementation to compare against.

@@

\section{Shrinking Data}\label{sec:shrinking}
In this section, we describe how to efficiently and generically shrink algebraic
data values.  Recall the basic idea behind the \ttp{shrink} method of the
\ttp{Arbitrary} class: generate a list of values, each of which is smaller than
the current counterexample.  Each of the new values generated may not bear any
relationship to the original counterexample other than being smaller.
SmartCheck pursues an approach that searches for smaller but structurally
similar counterexamples, as we make precise below.  Perhaps the most significant
difference between SmartCheck's reduction algorithm and QuickCheck shrink
implementations is that the latter is deterministic.  SmartCheck combines
counterexample search with shrinking.

@@

@@

@@

Unfortunately, no set of testing benchmarks exists over which to compare
different test-case generation and minimization approaches.  Therefore, we have
collected a small number of benchmarks, in addition to the more involved
case-studies described earlier in this section.  However, these are contrived
insofar as initial counterexamples for them are discovered quickly.

The benchmarks presented, in addition to the motivating example presented in
Section~\ref{sec:example}, compare standard SmartCheck against QuickCheck's
generic shrink implementation, which is, in general, as good or better than
hand-written shrink implementations.  The benchmarks are as follows:

@@

  \item \emph{Parser}, a parser/pretty-printer for a toy imperative language
    containing a parser bug that switches the arguments of an binary expression
    operator.
\end{itemize}

@@

@@

\multirow{2}{*}{Reverse}
  & QC & 2 & 0.002       & 0 & 0.002       & 2 & 0.003 \\\cline{2-8}
  & SC & 2 & \sci{4}{-4} & 0 & \sci{5}{-4} & 2 & \sci{7}{-4}\\\hline\hline

\multirow{2}{*}{Div0}
  & QC & 5  & 0.004 & 1 & 0.006 & 7  & 0.015 \\\cline{2-8}
  & SC & 5  & 0.001 & 0 & 0.001 & 5  & 0.001  \\\hline\hline

\multirow{2}{*}{Heap}
  & QC & 19 & \sci{9}{-4}  & 9 & 0.001 & 36 & 0.001 \\\cline{2-8}
  & SC & 7  & 0.006        & 2 & 0.002 & 10 & 0.010 \\\hline\hline

\multirow{2}{*}{Parser}
  & QC &  4   & 0.024 & 0  & 0.020 & 4 & 0.061 \\\cline{2-8}
  & SC &  6   & 0.116 & 2  & 0.083 & 9 & 0.230 \\\hline

@@

The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit
shrinking in general, since the original counterexamples are generally close to
being minimal.  Surprisingly, SmartCheck slightly outperforms QuickCheck in
efficiency.  The other three benchmarks have larger counterexamples, so the
benefit of shrinking is more pronounced.

SmartCheck finds smaller counterexamples in the \emph{Div0} and \emph{Heap}
benchmarks, while QuickCheck shrinking finds smaller counterexamples faster in
the \emph{Parser} example.  The \emph{Parser} example is one in which
SmartCheck's strategy of generating random values during counterexample
reduction is not a benefit, and so this is why it likely under performs a
generic \ttp{shrink} method.  Moreover, because SmartCheck uses QuickCheck as a
front-end, the two approaches are complementary: QuickCheck shrinking can be
used instead of or in combination with SmartCheck reduction.

@@

Zeller and Hildebrandt describe an application of greedy search to shrink
counterexamples they call ``delta-debugging'' (DD)~\cite{dd}.  The authors apply
their work to shrinking HTML inputs to crash Mozilla and shrinking C programs to
trigger a bug in GCC.  Subsequent generalizations are reported by Misherghi and
Su in which they perform greedy search on tree-structured data; they call their
approach hierarchical delta-debugging (HDD)~\cite{hdd}.
